The Effect of Oil Price on Field Production: Evidence from the Norwegian Continental Shelf

### 2015-03-31 mc_sim.rmd
update on paper

The function generates a starting year for field production from a half-normal distribution with a mean at 1970, but constrained to be between 1970 to 2000.  This is meant to reflect the common industry pattern that most significant finds are found early in a regions activity.

```{r}
gen_year<-function(size){
	#half-normal distribution with mean at 1970
	#constrained to be between 1970 and 2000
	range<-FALSE
	while(range==FALSE){
		year<-trunc(rnorm(1,mean=(1970), sd=10))
		ifelse(year>=1970 & year<=2000, range<-TRUE, range<-FALSE)
		}

@@

```{r}
add_uncertainty<-function(beta, fields){ 
	fields$prod<-fields$prod_shape*
		exp(beta*(fields$prices))*
			rlnorm(length(fields$prices), meanlog=0, sdlog=.10)
	return(fields)
}
```

The below function takes a regression formula and a data frame of field production data and runs a GAM regression, returning the coefficient on $prices$ variable.
```{r}
run_gam<-function(formula, fields){
	#runs GAM model and returns coefficient on price

@@

The below function takes a formula and a data frame of field production data and runs a GLM model, returning the estimated coefficient on the $prices$ variable. 
```{r}
run_glm<-function(formula, fields){	
	glm_sim<-lm(formula, data=fields)
	return(coefficients(glm_sim)["prices"])
}
```

@@

Now we create the simulated smoothed production profiles of the fields
```{r}
sim_fields<-apply(fields, 1, gen_production_shape)
sim_fields<-Reduce(rbind, sim_fields)
sim_fields<-merge(sim_fields, prices, by="year")
sim_fields<-sim_fields[order(sim_fields$name, sim_fields$year),]
````

Below we do a single run of generating production uncertainty, while setting the effect of prices to zero to visualize what the data looks like.  It looks reasonable.     
```{r ggplot, fig.width=8, fig.height=5, message=FALSE}
sim_fields_unc<-add_uncertainty(fields=sim_fields, beta=0)
ggplot(sim_fields_unc)+
 geom_line(aes(x=year, y=prod, color=factor(name)),alpha=.3, size=1) +
 guides(color=FALSE) +
 scale_color_grey(start=0, end=.3) +
 theme_bw() +
 labs(x="", y="Simulated Production from Fields")
 ```
###GLM Monte Carlo Experiment
Lets start with a monte-carlo regression using standard GLM, using a 4th-degree polynomial in order to try to control for the production profile of the fields.  
```{r}
formula_glm <- formula(I(log(prod)) ~ prod_years + I(prod_years^2) + I(prod_years^3) + 
	I(prod_years^4) + size + prices + year + I(year^2) + I(year^3))
```

@@

We transform the data in order to make it easier to visualize
```{r}
glm_mc_data<-data.frame(beta0=glm_mc_beta0, beta05=glm_mc_beta05)
glm_mc_data<-melt(glm_mc_data)
colnames(glm_mc_data)<-c("beta", "coefficient")
levels(glm_mc_data$beta)<-c("0", "0.05")
```

Below we plot the results in the form of empirical density functions of the results.  The vertical black lines represent the true values of beta.  Clearly, the GLM estimation has a significant negative bias.
```{r, fig.width=8, fig.height=5, message=FALSE}
ggplot(glm_mc_data, aes(x=coefficient, fill=beta)) +
geom_density(alpha=.5, size=0) +
geom_vline(aes(xintercept=c(0, 0.05))) +
scale_fill_grey() +
theme_bw() +
guides(fill=FALSE) +
xlab("beta_hat, estimated by GLM") +
ylab("Density")
```
### GAM Monte Carlo Experiment
Below we set the formula as a slightly simplified version of our preffered GAM model. I have removed the cost index as this was not taken into account when generating the data.  For simplicity I also do not take into account lagged price terms as the main point of this monte-carlo experiment is to show how the production profile can be effectively controlled for.

@@

Converting the data to a format for plotting
```{r}
gam_mc_data<-data.frame(beta0=gam_mc_beta0, beta05=gam_mc_beta05)
gam_mc_data<-melt(gam_mc_data)
colnames(gam_mc_data)<-c("beta", "coefficient")
levels(gam_mc_data$beta)<-c("0", "0.05")
levels(gam_mc_data)
```

Plotting the results, we see that the GAM model does a substantially better job of getting close to the "true" value.  
```{r, fig.width=8, fig.height=5, message=FALSE}
ggplot(gam_mc_data, aes(x=coefficient, fill=beta)) +
geom_density(alpha=.5, size=0) +
geom_vline(aes(xintercept=c(0, 0.05))) +
xlab("beta_hat, estimated by GAM") +
ylab("") +
scale_fill_grey() +
theme_bw()
```

>>>>

The function generates a starting year for field production from a half-normal distribution with a mean at 1970, but constrained to be between 1970 to 1990.  This is meant to reflect the common industry pattern that most significant finds are found early in a regions activity.

```{r}
gen_year<-function(sizexz){
	#half-normal distribution with mean at 1970
	#constrained to be between 1970 and 1990
	range<-FALSE
	while(range==FALSE){
		year<-trunc(rnorm(1,mean=(1970), sd=10))
		ifelse(year>=1970 & year<=2000, range<-TRUE, range<-FALSE)
		}

@@

```{r}
add_uncertainty<-function(beta, fields){ 
	fields$prod<-fields$prod_shape*
		exp(beta*(fields$prices))*
			rlnorm(length(fields$prices), meanlog=0, sdlog=.05)
	return(fields)
}
```

The below function takes a regression formula and a data frame of field production data and runs a GAM regression, returning the coefficient on $prices$ variable.

```{r}
run_gam<-function(formula, fields){
	#runs GAM model and returns coefficient on price

@@

The below function takes a formula and a data frame of field production data and runs a GLM model, returning the estimated coefficient on the $prices$ variable.

```{r}
run_glm<-function(formula, fields){	
	glm_sim<-lm(formula, data=fields)
	return(coefficients(glm_sim)["prices"])
}
```

@@

Now we create the simulated smoothed production profiles of the fields

```{r}
sim_fields<-apply(fields, 1, gen_production_shape)
sim_fields<-Reduce(rbind, sim_fields)
sim_fields<-merge(sim_fields, prices, by="year")
sim_fields<-sim_fields[order(sim_fields$name, sim_fields$year),]
````

Below we do a single run of generating production uncertainty, while setting the effect of prices to zero to visualize what the data looks like.  It looks reasonable.

```{r ggplot, fig.width=8, fig.height=5, message=FALSE}
sim_fields_unc<-add_uncertainty(fields=sim_fields, beta=0)
ggplot(sim_fields_unc)+
 geom_line(aes(x=year, y=prod, color=factor(name)),alpha=.3, size=1) +
 guides(color=FALSE) +
 scale_color_grey(start=0, end=.3) +
 theme_bw() +
 labs(x="", y="Simulated Production from Fields")
 ```

##GLM Monte Carlo Experiment

Lets start with a monte-carlo regression using standard GLM, using a 4th-degree polynomial in order to try to control for the production profile of the fields.

```{r}
formula_glm <- formula(I(log(prod)) ~ prod_years + I(prod_years^2) + I(prod_years^3) + 
	I(prod_years^4) + size + prices + year + I(year^2) + I(year^3))
```

@@

We transform the data in order to make it easier to visualize

```{r}
glm_mc_data<-data.frame(beta0=glm_mc_beta0, beta05=glm_mc_beta05)
glm_mc_data<-melt(glm_mc_data)
colnames(glm_mc_data)<-c("beta", "coefficient")
levels(glm_mc_data$beta)<-c("0", "0.05")
```

Below we plot the results in the form of empirical density functions of the results.  The vertical black lines represent the true values of beta.  Clearly, the GLM estimation has a significant negative bias.

```{r, fig.width=8, fig.height=5, message=FALSE}
ggplot(glm_mc_data, aes(x=coefficient, fill=beta)) +
geom_density(alpha=.5, size=0) +
geom_vline(aes(xintercept=c(0, 0.05))) +
scale_fill_grey() +
theme_bw() +
guides(fill=FALSE) +
xlab("beta_hat, estimated by GLM") +
ylab("Density")
```

## GAM Monte Carlo Experiment

Below we set the formula as a slightly simplified version of our preffered GAM model. I have removed the cost index as this was not taken into account when generating the data.  For simplicity I also do not take into account lagged price terms as the main point of this monte-carlo experiment is to show how the production profile can be effectively controlled for.

@@

Converting the data to a format for plotting

```{r}
gam_mc_data<-data.frame(beta0=gam_mc_beta0, beta05=gam_mc_beta05)
gam_mc_data<-melt(gam_mc_data)
colnames(gam_mc_data)<-c("beta", "coefficient")
levels(gam_mc_data$beta)<-c("0", "0.05")
levels(gam_mc_data)
```

Plotting the results, we see that the GAM model does a substantially better job of getting close to the "true" value.

```{r, fig.width=8, fig.height=5, message=FALSE}
ggplot(gam_mc_data, aes(x=coefficient, fill=beta)) +
geom_density(alpha=.5, size=0) +
geom_vline(aes(xintercept=c(0, 0.05))) +
xlab("beta_hat, estimated by GAM") +
ylab("") +
scale_fill_grey() +
theme_bw()
```

### 2015-03-31 oil_prices_old.tex
update on paper

This estimate is heavily biased downwards due to the spurious correlation between the field production profiles and the autocorrelated and non-stationary time series of oil prices.  A more formal analysis of the bias can be found in the online appendix, \url{jmaurit.github.io#oil_price_sim}, where I simulated data of oil field production in a monte carlo experiment.  However, the basic problem is that the parametric representation is not flexible enough to control for the production profiles of the fields.   Instead, a more flexible estimation of the production profile is needed.

>>>>

This estimate is heavily biased downwards due to the spurious correlation between the field production profiles and the autocorrelated and non-stationary time series of oil prices.  A more formal analysis of the bias can be found in the on-line appendix, \url{jmaurit.github.io#oil_price_sim}, where I simulated data of oil field production in a monte carlo experiment.  However, the basic problem is that the parametric representation is not flexible enough to control for the production profiles of the fields.   Instead, a more flexible estimation of the production profile is needed.
